---
title: "Assignment 4: Capital Bikeshare"
author: "Ruihan He"
date: "`9-16-2025`"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    dev: pdf
    highlight: tango
  html_document:
    theme: default
    self_contained: true
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    smart: true
    dev: svg
---

```{r setup, include = FALSE, echo=FALSE}
# DO NOT ALTER THIS CHUNK
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.width = 5,
  fig.asp = 0.618,
  dpi = 120,
  fig.align = "center",
  cache = FALSE
)
# Load required packages
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(lubridate))
suppressPackageStartupMessages(library(geosphere))
bikeshare <- readRDS("bikeshare_data.RDS")
```


## Exercise 1
start_lat  end_lat
rideable_type member_causal
Each row = one bike trip.

## Exercise 2
```{r}
bikeshare <- bikeshare %>%
  mutate(
    duration = ended_at - started_at
  )
```


## Exercise 3
```{r}
bikeshare %>%
  ggplot() +
  geom_histogram(mapping = aes(x = duration)) +
  labs(title = "Distribution of Capital Bikeshare Trip Durations", x = "Duration (seconds)")
```
The histogram shows that most Capital Bikeshare trips are short, but there is a long tail of longer trips, resulting in a right-skewed distribution.


## Exercise 4
```{r}
library(ggplot2)
library(dplyr)

bikeshare %>%
  ggplot() +
  geom_boxplot(aes(y = duration)) +
  labs(
    title = "Box Plot of Capital Bikeshare Trip Durations",
    y = "Duration (seconds)"
  )
```
Box plot code was adapted from https://r-graph-gallery.com/boxplot.html. 
I modified it to use the bikeshare dataframe and the duration variable.

Center: The middle line of the box shows the median duration.
Skewness: The box plot is right-skewed, just like the histogram, because there are some very long trips.
Comparison to histogram:
Both show right skew.
Box plot highlights the median, quartiles, and extreme values more clearly.


## Exercise 5



## Exercise 6



## Exercise 7



## Exercise 8



## Exercise 9



## Exercise 10



## Exercise 11




## Academic Integrity statement




