---
title: "Lab 2: Exploration by visualization: the streaming movies dataset"
author: "Ruihan He"
date: "9-18-2025"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    highlight: tango
    toc: false
    df_print: kable
    fig_caption: no
    number_sections: no
    dev: pdf
---

```{r setup, include = FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
# Set knitr options
knitr::opts_chunk$set(
  echo = TRUE, eval = TRUE, fig.width = 5,
  fig.asp = 0.618, out.width = "80%", dpi = 120,
  fig.align = "center", cache = FALSE
)

# Load packages
suppressPackageStartupMessages(library(tidyverse))

# Load dataset
streaming <- read_csv("streaming.csv")
```

* * *

## Visualization by example

### Exercise 1

```{r}
qplot(x = im_db, binwidth = 0.5, data = streaming)
```
The histogram simplifies the raw im_db data into an easily interpretable visual, highlighting where most ratings concentrate and how they spread across possible values.


### Exercise 2

```{r}
qplot(
  x = im_db, 
  binwidth = 0.5, 
  fill = main_language,
  data = streaming
  )
```
Adding fill = main_language to the histogram code created a stacked histogram where each bar (representing a bin of IMDb ratings) is divided into colored segments based on the main_language variable. 
From the histogram, it is clear that most movies in the database are in English.


### Exercise 3




### Exercise 4





### Exercise 5




### Exercise 6





### Exercise 7





### Exercise 8

  


### Exercise 9



