---
title: "Lab 2: Exploration by visualization: the streaming movies dataset"
author: "Ruihan He"
date: "9-18-2025"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    highlight: tango
    toc: false
    df_print: kable
    fig_caption: no
    number_sections: no
    dev: pdf
---

```{r setup, include = FALSE}
# DO NOT ALTER CODE IN THIS CHUNK
# Set knitr options
knitr::opts_chunk$set(
  echo = TRUE, eval = TRUE, fig.width = 5,
  fig.asp = 0.618, out.width = "80%", dpi = 120,
  fig.align = "center", cache = FALSE
)

# Load packages
suppressPackageStartupMessages(library(tidyverse))

# Load dataset
streaming <- read_csv("streaming.csv")
```

* * *

## Visualization by example

### Exercise 1

```{r}
qplot(x = im_db, binwidth = 0.5, data = streaming)
```
The histogram simplifies the raw im_db data into an easily interpretable visual, highlighting where most ratings concentrate and how they spread across possible values.


### Exercise 2

```{r}
qplot(
  x = im_db, 
  binwidth = 0.5, 
  fill = main_language,
  data = streaming
  )
```
Adding fill = main_language to the histogram code created a stacked histogram where each bar (representing a bin of IMDb ratings) is divided into colored segments based on the main_language variable. 
From the histogram, it is clear that most movies in the database are in English.


### Exercise 3
The distribution of English appears roughly symmetric or slightly left-skewed, with most ratings clustered in the mid-range. Its modality is unimodal, centered around 6–7.
The distribution of Foreign is slightly right-skewed, with a concentration of ratings in the 6–8 range and a noticeable number of films rated above 8. Its modality is unimodal,a peak around 7–7.5.
English films dominate the dataset numerically, but Foreign films tend to be rated more favorably on average.
Yes. Because it will be influenced by factors such as selection bias, cultural perception and personal preferences.


### Exercise 4
```{r}
qplot(
  x = im_db, 
  binwidth = 0.5, 
  facets = ~ age,
  data = streaming
  )
```

Facets: 5
Variable faceted over: Age rating category
Facet with most movies: “18+”


### Exercise 5
```{r}
qplot(x = rotten_tomatoes, y = im_db, data = streaming)
```
As Rotten Tomatoes ratings increase, IMDb ratings also tend to increase in a roughly linear pattern. Movies with very low Rotten Tomatoes scores generally have lower IMDb scores, while movies with very high Rotten Tomatoes scores tend to have higher IMDb scores. This suggests that the two ratings systems are positively associated.


### Exercise 6
```{r}
library(ggplot2)

ggplot(streaming, aes(x = rotten_tomatoes, y = im_db, color = main_language)) +
  geom_point()
```
The basic association looks similar for both English and foreign-language movies, but the spread and density of the points differ.


### Exercise 7
```{r}
library(ggplot2)

ggplot(streaming, aes(x = rotten_tomatoes, y = im_db)) +
  geom_point() +
  facet_wrap(~ age)   # or facet_wrap(~age_rating) depending on your variable name

```
The faceted view doesn’t show a fundamentally different relationship from the colored view.




### Exercise 8

  


### Exercise 9



