---
title: "Assignment 2: Visualization"
author: Ruihan He
date: "`9-3-2025`"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    dev: pdf
    highlight: tango
  html_document:
    theme: default
    self_contained: true
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    smart: true
    dev: svg
---

```{r setup, include = FALSE}
# DO NOT ALTER THIS CHUNK
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.width = 5,
  fig.asp = 0.618,
  out.width = "70%",
  dpi = 120,
  fig.align = "center",
  cache = FALSE,
  warning = FALSE,
  message = FALSE
)
```

## Exercise 1

```{r}
library(palmerpenguins)
data(penguins) 
# View the structure of the dataset 
str(penguins)
```

Number of rows and columns: 344 rows, 8 columns.

Each row represents an observation of a penguin.

Categorical variables: species, island, sex.

Continuous variables: bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g.

Variable that could be categorical or continuous: year.

Species included: Adelie, Chinstrap, Gentoo.

## Exercise 2

```{r}
library(ggplot2)
ggplot(penguins, aes(x = flipper_length_mm)) + geom_histogram()
```

The x-axis shows flipper length (mm).

The numbers on the axis represent values of flipper length.

The distribution is bimodal.

## Exercise 3

```{r}
ggplot(penguins, aes(x = flipper_length_mm)) + geom_histogram(bins = 10)  # fewer bins
ggplot(penguins, aes(x = flipper_length_mm)) + geom_histogram(bins = 40)  # more bins
```

With fewer bins → the histogram looks smoother but shows less detail.

With more bins → the histogram looks noisier but reveals more detail.

Increasing the number of bins makes the two-peak pattern easier to see.

## Exercise 4

```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + geom_point()
```

The y-axis represents body mass (g).

There is a positive relationship between flipper length and body mass.

The relationship is roughly linear.

## Exercise 5

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g)) + geom_point()
```

The correlation between bill length and body mass is weaker than the correlation between flipper length and body mass.

The points are more scattered, and the trend is less obvious.

## Exercise 6

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g)) + geom_point() + 
geom_smooth(method = "lm", se = FALSE)
```

Adding a trend line makes the overall pattern clearer.

However, it may also hide differences among species.

## Exercise 7

```{r}
ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species)) + geom_point()
```

Coloring by species makes the relationship easier to understand:

Gentoo: longer bills and heavier body mass.

Adelie: shorter bills and medium body mass.

Chinstrap: long bills but lighter body mass.

Color highlights species-specific patterns that are harder to see otherwise.
