---
title: "Assignment 2: Visualization"
author: Ruihan He
date: "`9-3-2025`"
documentclass: article
geometry: margin=1in
fontsize: 11pt
output:
  pdf_document:
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    dev: pdf
    highlight: tango
  html_document:
    theme: default
    self_contained: true
    toc: false
    df_print: kable
    fig_caption: false
    number_sections: false
    smart: true
    dev: svg
---

```{r setup, include = FALSE}
# DO NOT ALTER THIS CHUNK
knitr::opts_chunk$set(
  echo = TRUE,
  eval = TRUE,
  fig.width = 5,
  fig.asp = 0.618,
  out.width = "70%",
  dpi = 120,
  fig.align = "center",
  cache = FALSE,
  warning = FALSE,
  message = FALSE
)
```

## Exercise 1
```{r}
library(palmerpenguins)
data(penguins) 
# View the structure of the dataset 
str(penguins)
```

Number of rows and columns: 344 rows, 8 columns.
Each row represents an observation of a penguin.
Categorical variables: species, island, sex.
Continuous variables: bill_length_mm, bill_depth_mm, flipper_length_mm, body_mass_g.
Variable that could be categorical or continuous: year.
Species included: Adelie, Chinstrap, Gentoo.

## Exercise 2
```{r}
library(ggplot2)
ggplot(penguins, aes(x = flipper_length_mm)) + geom_histogram()
```

## Exercise 3
```{r}
ggplot(penguins, aes(x = flipper_length_mm)) + geom_histogram(bins = 10)  # fewer bins
ggplot(penguins, aes(x = flipper_length_mm)) + geom_histogram(bins = 40)  # more bins
```

## Exercise 4
```{r}
ggplot(penguins, aes(x = flipper_length_mm, y = body_mass_g)) + geom_point()
```

## Exercise 5
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g)) + geom_point()
```


## Exercise 6
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g)) + geom_point() +
geom_smooth(method = "lm", se = FALSE)
```

## Exercise 7
```{r}
ggplot(penguins, aes(x = bill_length_mm, y = body_mass_g, color = species)) + geom_point()
```

